A thesis draft

### 2018-11-20 res/sections/08-Background.tex
Add citations and solve TODOs into 5g background

\section{The upcoming connectivity standard: 5G}
The continuous innovation in the mobile network connectivity is leading to the
creation of a new standard, the 5G, which is estimated to arrive to the market
consumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
alliance, a group composed by the major players in the field of mobile
connectivity, the 5G aims to offer not only at the end-user a new way to browse 
the web, download and watch interactive content but also to create an ad-hoc 
solution for Machine-to-Machine (M2M) data traffic, which is increasing more 
than ever thanks to the spreading of IoT devices, which sensors need to 
continuously send data to servers/data-centers. Relatively to LTE\todo{Explain 
acronym}, 5G points to have data rates $10$ times better, with $10$ times 
smaller end-to-end latency and an increased connection density by $100$ 
times\todo{CITE}.

@@

Part of the new requirements can be satisfied using a large radio spectrum with
higher frequencies. The utilization of higher frequencies, though, mean that the
radio signals can be easily disrupted by physical objects, like buildings and
many geographical elements (such as hills and mountains), clashing with the
expectation of an ever-reachable connectivity. It is here where virtualization
plays an important role. In fact, the re-design of some network components today
existing via hardware can transform a monolithic networking approach to a 
modular one, exploiting the flexibility that Virtual Network Functions 
(VNFs)\todo{CITE VNF-P: A Model for Efficient Placement of Virtualized Network 
Functions} can offer thanks to virtualization, closing the gap to the use-case 
fulfillment defined by the NGMN alliance\todo{CITE: NGMN View on 5G 
architecture} which require a decreased time to set up and deploy network 
services (specifically, from 90 hours to 90 minutes\todo{Search this 
requirement, check 90h 'cause I'm not sure about it}).

@@

With the constraints placed on the requirements formulated by the NGMN alliance,
5G envisage a multi-layered architecture, based on three main layers:\todo{CITE:
  NGMN View on 5G architecture}
\begin{itemize}
\item \textbf{infrastructure resource layer}: physical resources that are 
exposed via a virtualized interface, and that can be monitored using specific 
APIs
\item \textbf{business enablement layer}: where a library of functions and
  deployment is contained, and its configuration is accessible via APIs
\item \textbf{business application layer}: layer that contains specific
  applications and services of the operator
\end{itemize}

\todo{Add three-layer image here?}

@@

\subsubsection{Network slicing}
The role of a ``network slice'' in a 5G architecture is to specifically handle 
the Control-plane\todo{Explain what a control plane is?} of a particular 
service (e.g. smartphones traffic, autonomous driving, massive IoT), deploying 
resources in a manner that assure the required latency, security and 
reliability. While some very peculiar legacy services could require specific 
hardware, the common resources between services could be shared in a 
virtualized way, providing auto-scaling capabilities in services that are under 
heavy network pressure.

@@

% \section{Technologies used}

>>>>

\section{The upcoming connectivity standard: 5G}
The continuous innovation in the mobile network connectivity is leading to the
creation of a new standard, the 5G, which is estimated to arrive to the market
consumer in 2020~\cite{iwamura2015ngmn}. Lead by the Next Generation Mobile
Network (NGMN) alliance, a group composed by the major players in the field of
mobile connectivity, the 5G aims to offer not only at the end-user a new way to
browse the web, download and watch interactive content but also to create an
ad-hoc solution for Machine-to-Machine (M2M) data traffic, which is increasing
more than ever thanks to the spreading of IoT devices, which sensors need to 
continuously send data to servers/data-centers. Relatively to \emph{Long Term
Evolution} (LTE), 5G points to have data rates $10$ times better, with $10$
times smaller end-to-end latency and an increased connection density by $100$ 
times~\cite{alliance20155g}.

@@

Part of the new requirements can be satisfied using a large radio spectrum with
higher frequencies. The utilization of higher frequencies, though, mean that the
radio signals can be easily disrupted by physical objects, like buildings and
many geographical elements (such as hills and mountains), clashing with the
expectation of an ever-reachable connectivity. It is here where virtualization
plays an important role. In fact, the re-design of some network components today
existing via hardware can transform a monolithic networking approach to a 
modular one, exploiting the flexibility that Virtual Network Functions 
(VNFs)\cite{moens2014vnf} can offer thanks to virtualization, closing the gap to
the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
which require a decreased time to set up and deploy network services
(specifically, from 90 hours to 90 minutes~\cite{networld20202014role}).

@@

With the constraints placed on the requirements formulated by the NGMN alliance,
5G envisage a multi-layered architecture, based on three main
layers~\cite{alliance20155g}:
\begin{itemize}
\item \textbf{infrastructure resource layer}: physical resources that are 
exposed via a virtualized interface, and that can be monitored using specific 
APIs
\item \textbf{business enablement layer}: where a library of functions and
  deployment is contained, and its configuration is accessible via APIs
\item \textbf{business application layer}: layer that contains specific
  applications and services of the operator
\end{itemize}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.55]{5garchitecture}
  \caption{5G Architecture~\cite{alliance20155g}}
  \label{chap:background:img:5garchitecture}
\end{figure}

@@

\subsubsection{Network slicing}
The role of a \emph{network slice} in a 5G architecture is to specifically
handle the Control-plane of a particular service (e.g. smartphones traffic,
autonomous driving, massive IoT), deploying resources in a manner that assure
the required latency, security and reliability. While some very peculiar legacy
services could require specific hardware, the common resources between services
could be shared in a virtualized way, providing auto-scaling capabilities in
services that are under heavy network pressure.

@@
